---
title: "Problem Set 7"
output: html_document
date: "2023-10-20"
---

# Question 1e

```{r}
mean <- 0
n <- 100
xbar <- 0.5
sd <- 2.323
Z <- (xbar - mean)/(sd/sqrt(n))
Z
```
```{r}
pnorm(Z)
```
```{r}
1 - pnorm(Z)
```


# Question 2b

Given: 
n = 1000
mean = 9.06,
variance = 565.2476
Standard deviation = ?
Standard error = ?

```{r}
mean <- 9.06
sd <- sqrt(565.2476)
se <- sd/sqrt(1000)
```

95% confidence for the mean number of citations
alpha = 0.05
alpha/2 = 0.025

```{r}
upper <- mean + qnorm(0.975,0,1) * se
lower <- mean - qnorm(0.975,0,1) * se

print(paste("The approximate 95% confidence interval for the mean number of citations is" ,upper ,"and" ,lower))
```

# Question 2c

Given that 460 out of 1000 articles don't have citations
This question can be turned into a bernoulli trial
Probability that a article has no citation p = 460/1000 = 0.46
With 1000 trials, xbar ~ equal to p
We know that sd = sqrt(xbar*(1-xbar))
```{r}
xbar <- 0.46
n <- 1000                               
sd <- sqrt(xbar*(1-xbar)) 
se <- sd/sqrt(n)
print(paste("standard deviation = " ,sd)) 
print(paste("standard error is = " , se))
```
```{r}
upper <- xbar + qnorm(0.975) * se
lower <- xbar- qnorm(0.975) * se

print(paste("The 95% confidence interval of journal with no citation is",upper , "and" ,lower))
```

# Question 3

Reading the data

```{r}
data <- read.table("ANES2016.txt",header = TRUE)
View(data)
nrow(data)
```

Plotting the distribution of the data

```{r}
library(ggplot2)
ggplot(data, aes(x = Thermometer)) + geom_histogram()
```

Plotting Trump alone

```{r}
trump <- subset(data,Candidate == "Trump")
ggplot(trump, aes(x = Thermometer)) + geom_histogram()
```

Plotting Clinton alone

```{r}
clinton <- subset(data,Candidate == "Clinton")
ggplot(clinton, aes(x = Thermometer)) + geom_histogram()
```

Plotting Sanders alone

```{r}
sanders <- subset(data,Candidate == "Sanders")
ggplot(sanders, aes(x = Thermometer)) +geom_histogram()
```

# Question 3b

```{r}
temp100 <- subset(data, Thermometer <= 100)

trump <- subset(temp100, Candidate == "Trump")
meanTrump <- mean(trump$Thermometer)
sdTrump = sd(trump$Thermometer)

clinton <- subset(temp100, Candidate == "Clinton")
meanClinton <- mean(clinton$Thermometer)
sdClinton <- sd(clinton$Thermometer)

sanders <- subset(temp100, Candidate == "Sanders")
meanSanders <- mean(sanders$Thermometer)
sdSanders <- sd(sanders$Thermometer)

print(paste("Trump's Mean = ", meanTrump," Standard Deviation = ", sdTrump ))
print(paste("Clinton's Mean = ", meanClinton," Standard Deviation = ", sdClinton ))
print(paste("Sanders' Mean = ", meanSanders," Standard Deviation = ", sdSanders ))
```

# Question 3c

Find 99% confidence intervals for:

i. Trump’s mean feeling thermometer score if all adult U.S. citizens were asked in January 2016.
alpha = 0.01; alpha/2 = 0.005

```{r}
seTrump <- sd(trump$Thermometer)/sqrt(1197)
upper <- meanTrump + qnorm(0.995) * seTrump    
lower <- meanTrump - qnorm(0.995) * seTrump
print(paste("99% confidence interval of Trumps mean feeling thermometer of", meanTrump," is", upper, "and", lower))
```

ii. Clinton’s mean feeling thermometer score if all adult U.S. citizens were asked in January 2016.

```{r}
seClint <- sd(clinton$Thermometer)/sqrt(1199)
upper <- meanClinton + qnorm(0.995) * seTrump        
lower <- meanClinton - qnorm(0.995) * seTrump
print(paste("99% confidence interval of Clinton mean feeling thermometer of", meanClinton," is", upper, "and", lower))
```

iii. Sanders’ mean feeling thermometer score if all adult U.S. citizens were asked in January 2016.

```{r}
seSanders <- sd(sanders$Thermometer)/sqrt(1192)
upper <- meanSanders + qnorm(0.995) * seTrump   
lower <- meanSanders - qnorm(0.995) * seTrump
print(paste("99% confidence interval of Sanders mean feeling thermometer of", meanSanders," is", upper, "and", lower))
```

# Question 4

Given, SD = 6
We know n = (2 * q * SD / L)^2

Given interval length = 2 feet
Q value for 99% interval = 
```{r}
qnorm(0.995)
```

Solving for n =
```{r}
n = (2 * qnorm(0.995) * 6/2)^2
n
```

Therefore, n = 238.85 ~ 239


# Question 5a

This can be considerd a bernoulli trial
A person either supports or does not support same-sex marriage
If 63% of 1009 support, then p ~ 0.63
In a bernoulli random variable, we can assume p to be same as the expected value (xbar = 0.63) 

```{r}
xbar <- 0.63
sd <- sqrt(xbar *(1-xbar))
se <- sd / sqrt(1009)
se
```

```{r}
upper <- xbar + qnorm(0.975) * se
lower <- xbar - qnorm(0.975) * se
print(paste("95% confidence interval for the percentage of all US adults who supports same sex marriage is", upper, "and" , lower))
```

# Question 5b


```{r}
confidence_level <- 0.95
length_confidence_interval <- 2

z <- qnorm((1 + confidence_level) / 2)
n <- (2 * z * sd / length_confidence_interval) ^ 2

print(paste("The sample size required will be = ", n))
```






