---
title: "In-class 10/24/23"
author: "S520"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Depression data:
```{r}
Before <- c(1.83, 0.50, 1.62, 2.48, 1.68, 1.88, 1.55, 3.06, 1.30)
After <- c(0.88, 0.65, 0.60, 2.05, 1.06, 1.29, 1.06, 3.14, 1.29)
plot(Before, After)
```
```{r}
Changes <- After - Before
Changes
```

Let mu be the expected change in HSF score after taking the drug.

H0 : mu >= 0

H1 : mu < 0

This is a one-tailed test.

We have an extremely small sample. Can we do a t-test?

This would assume the data comes from a Normal distribution.

```{r}
qqnorm(Changes)
```

Leap of faith: Assume that the population is Normal.

Calculate your t-statistic:
```{r}
xbar <- mean(Changes)
s <- sd(Changes)
n <- length(Changes)
c(xbar, s, n)
```
```{r}
t.stat <- (xbar - 0) / (s / sqrt(n))
t.stat
```

Compare this to the left tail of a t-distribution with n-1 = 8 degrees of freedom:
```{r}
pt(t.stat, df = n - 1)
```

The P-value is less than 1%, which is quite small.

Find a 95% CI:
```{r}
q <- qt(.975, df = 8)
q
xbar - q * s / sqrt(n)
xbar + q * s / sqrt(n)
```

Our 95% CI says that the expected change (mu) in HSF score is a reduction between -0.75 and -0.10 units.

We have evidence that mu is less than zero. On average, the HSF score is lower after the treatment. However, this isn't necessarily because of the drug -- it could be that e.g. score "naturally" go down with time.



### Etrsucan and Italian skulls

```{r}
skulls <- scan("https://mtrosset.pages.iu.edu/StatInfeR/Data/skulls.dat")
```

You could split this vector into Etruscans and Italians:
```{r}
etruscans <- skulls[1:84]
italians <- skulls[85:154]
mean(etruscans)
mean(italians)
```

Or you could create a data frame that has a categorical variable that's either "Etruscan" or "Italian":
```{r}
pop <- c(rep("Etruscan", 84), rep("Italian", 70))
skulls.df <- data.frame(skulls, pop)
skulls.means <- aggregate(skulls ~ pop, FUN = mean, data = skulls.df)
skulls.means
```

Let Delta be the population mean breadth of Etruscan skulls minus the population mean breadth of Italian skulls.

Find a point estimate, Delta.hat:
```{r}
skulls.means$skulls
Delta.hat <- skulls.means$skulls[1] - skulls.means$skulls[2]
Delta.hat
```

Test the hypotheses (two-sample, two-tailed):

H0 : Delta = 0

H1 : Delta is not 0

Two options for the test:

- *Student's* two-sample t-test: IID samples from Normal pops with equal population variances.

- *Welch's* two-sample t-test: IID samples from Normal pops; variances may be equal or unequal.

Student's: Check Normality:
```{r}
library(ggplot2)
ggplot(skulls.df, aes(sample = skulls)) +
  stat_qq() + facet_grid(~pop)
```

Approx. Normality seems like a reasonable assumption.

Check equal variances:
```{r}
skulls.vars <- aggregate(skulls ~ pop, FUN = var, data = skulls.df)
skulls.vars
```

close enough.

Let's do Student's test:

Pooled sample variance:
```{r}
v1 <- skulls.vars$skulls[1]
v2 <- skulls.vars$skulls[2]
n1 <- 84
n2 <- 70
sp2 <- ((n1 - 1) * v1 + (n2 - 1) * v2) / (n1 + n2 - 2)
sp2
```

Student's t-statistic:
```{r}
t.stud <- (Delta.hat - 0) / sqrt(sp2 * (1/n1 + 1/n2))
t.stud
```

Compare to a t-distribution with n1 + n2 - 2 degrees of freedom:
```{r}
2 * (1 - pt(abs(t.stud), df = n1 + n2 - 2))
```

We reject the null that the pop means are equal. On average, Etruscan skulls are broader than Italian skulls.


Welch's two-sample test:

Find the standard error of Delta hat:
```{r}
std.error <- sqrt(v1 / n1 + v2 / n2)
std.error
```

Find Welch's t-statistic:
```{r}
t.welch <- Delta.hat / std.error
t.welch
```

Find the DF:
```{r}
DF <- (v1/n1 + v2/n2)^2 /
  ((v1/n1)^2/(n1 - 1) + (v2/n2)^2/(n2 - 1))
DF
```

Find the P-value:
```{r}
2 * (1 - pt(abs(t.welch), df = DF))
```

Same conclusion as Student's.

We should also find a 95% CI for Delta:
```{r}
q <- qt(.975, df = DF)
q
```
```{r}
Delta.hat - qt(.975, df = DF) * std.error
Delta.hat + qt(.975, df = DF) * std.error
```

We're 95% confident that the difference between av. Etruscan skull breadth and av. Italian skull breadth is 9 to 13 mm (Etruscans are broader.)


Easy way:
```{r}
# Welch's
t.test(skulls ~ pop, data = skulls.df)
```

```{r}
# Student's
t.test(skulls ~ pop, var.equal = TRUE, data = skulls.df)
```







