---
title: "Problem Set 12"
output: html_document
date: "2023-12-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question - 1: Trosset chapter 15.7 exercise 5, parts (a), (b), and (c)

### Input data
```{r}
sister <- c(69, 64, 65, 63, 65, 62, 65, 64, 66, 59, 62)
brother <- c(71, 68, 66, 67, 70, 71, 70, 73, 72, 65, 66)

length(sister)
length(brother)
```

### (a) sample coefficient of determination: the sample coefficient of determination tells how much of the variation in y is “explained” by the linear model. In simple linear regression, it’s literally just the square of the correlation between the 2 variables (sister and brother variables)

```{r}
cor(sister, brother)^2
```
Alternatively solution:
```{r}
# Linear regression
fit <- lm(brother ~ sister)

# R-squared
rsquared <- summary(fit)$r.squared
rsquared
```
### (b) To test whether knowing a sister's height helps predict her brother's height, we can perform a hypothesis test. The null hypothesis (H0) is that the slope (β1) is zero, indicating no relationship.
```{r}
summary(fit)
```
### From the linear model summary, we see that the P-value is 0.0744 which is not less than the given alpha (0.05). Therefore we fail to reject the null hypothesis and there is not enough evidence to conclude that knowing a sister's height helps predict her brother's height.

### (c) 0.90-level confidence interval for the slope of the population regression line for predicting y from x.
```{r}
conf_interval <- confint(fit, level = 0.90)
conf_interval
```
### The 0.90 level confidence interval is 0.054 - 1.1278.

